---
title: "Using depguard in Kaggle, Colab, and on your desktop"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using depguard in Kaggle, Colab, and on your desktop}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## The problem

Hosted, ephemeral R notebooks (Kaggle, Colab, Binder, and similar) ship with
a large pre-installed set of packages at fixed versions. When you install an
additional package, its dependencies can silently upgrade or downgrade a
package that is already loaded elsewhere in your session, breaking code
further down the notebook with no install-time error. The same thing happens
on a normal desktop whenever you install into a library you share between
projects.

Tools like `renv` solve this well for projects you fully own and can persist,
but assume you can write lockfiles and restore a private library. That often
doesn't fit a throwaway notebook, or a quick script. `depguard` fills the
narrower gap: lightweight checks that run entirely from locally installed
metadata, so they work offline.

## What is the environment?

```{r}
library(depguard)
dep_env()
```

`dep_env()` detects Kaggle, Colab, Binder, RStudio and plain desktop
sessions, and reports which libraries are writable. `depguard` uses it to
word its advice (for example, how to restart the session).

## Workflow 1: snapshot before, diff after

Run this before installing anything new:

```{r}
snap <- dep_snapshot(path = "depguard_snapshot.rds")
```

Install what you need as usual:

```{r}
install.packages("someNewPackage")
```

Then check what changed:

```{r}
dep_diff(snap)
```

The comparison is made on the versions *on disk*. R keeps running the old
version of an already-loaded package until you restart, so a package that
changed and is still loaded is flagged `risk = "high"` and `session_stale`:
code you already ran used one version, and code run after a restart will use
another.

In hosted notebooks you normally have to restart the session after installing.
Saving the snapshot with `path` lets you compare *after* the restart:

```{r}
dep_diff("depguard_snapshot.rds")
```

## Workflow 2: declare a manifest

If you know what your notebook needs up front, declare it once:

```{r}
dep_manifest(
  dplyr = ">= 1.1.4",   # at least this version
  ggplot2 = "3.5.0",    # a bare version also means "at least"
  quantmod = "== 0.4.26" # exactly this version
)
```

Or pin whatever you are using right now:

```{r}
library(dplyr)
dep_manifest_freeze(path = "depguard.txt")
```

Any file extension other than `.rds` produces a plain-text manifest that is
easy to read, edit and commit:

```
# depguard manifest: package (operator version)
dplyr (== 1.1.4)
```

Then, any time later:

```{r}
dep_check(path = "depguard.txt")
```

This walks the dependency tree of each declared package using only locally
installed metadata and compares the version constraints that the packages
declare on one another (for example, `cli (>= 3.4.0)`) with what is actually
installed. It reports `missing`, `mismatch`, and `restart` (the version on disk
is right but the session is still running an older one), along with which
package required each dependency. In scripts and CI, use
`dep_check(stop_on_problem = TRUE)` to fail on any problem.

## Workflow 3: one-shot health check

At the top of a notebook:

```{r}
dep_healthcheck()
```

This describes the environment, warns about packages hidden by another
library, and then checks a manifest if one exists, or captures a baseline
snapshot otherwise.

## "I upgraded it but nothing changed"

R loads the first copy of a package it finds on the library path. A stale copy
in a library earlier on the path hides the one you just installed.

```{r}
dep_libraries()
```

## Fixing a conflict

If a specific package got silently bumped and broke something, you can roll
just that package back to a specific version, and preview it first:

```{r}
dep_fix("stringr", "1.5.0", dry_run = TRUE)
dep_fix("stringr", "1.5.0")
```

`dep_fix()` uses `pak` or `remotes` when installed and otherwise downloads the
source from the CRAN archive using only base R. It checks the result and tells
you if you need to restart, or if another library is still hiding the rolled-back
copy.

This performs a single-package rollback only; it does not resolve cascading
conflicts the rollback might introduce elsewhere. For full dependency
resolution, use `renv::restore()` or `pak`'s solver.
