---
title: "Using depguard in Kaggle, Colab, and other sandboxed notebooks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using depguard in Kaggle, Colab, and other sandboxed notebooks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## The problem

Hosted, ephemeral R notebooks (Kaggle, Colab, Binder, and similar) ship with
a large pre-installed set of packages at fixed versions. When you install an
additional package, its dependencies can silently upgrade or downgrade a
package that is already loaded elsewhere in your session -- breaking code
further down the notebook with no install-time error.

Tools like `renv` solve this well for projects you fully own and can persist,
but assume you can write lockfiles and restore a private library. That
often doesn't fit a throwaway, read-mostly notebook session. `depguard`
fills the narrower gap: lightweight, local-first checks that work without
lockfile ownership.

## Workflow 1: snapshot before, diff after

Run this before installing anything new:

```{r}
library(depguard)

snap <- dep_snapshot()
```

Install what you need as usual:

```{r}
install.packages("someNewPackage")
```

Then check what changed:

```{r}
dep_diff(snap)
```

Packages that changed *and* are still loaded in your session are flagged
`risk = "high"` -- these are the ones most likely to break code you've
already run.

## Workflow 2: declare a manifest

If you know what your notebook needs up front, declare it once:

```{r}
dep_manifest(
  dplyr = "1.1.4",
  ggplot2 = "3.5.0",
  quantmod = "0.4.26"
)
```

Then, any time later in the session (or at the top of a rerun):

```{r}
dep_check()
```

This walks the transitive dependency tree of each declared package using
only locally installed metadata (no network calls by default) and reports
missing packages, version mismatches, and which top-level package pulled
in each transitive dependency.

## Workflow 3: one-shot health check

If you just want a sensible default at the top of a notebook:

```{r}
dep_healthcheck()
```

This checks against a manifest if one exists, or falls back to capturing a
baseline snapshot you can diff against later.

## Fixing a conflict

If a specific package got silently bumped and broke something, you can roll
just that package back:

<<<<<<< HEAD
```{r , eval = FALSE}
=======
```{r, eval = FALSE}
>>>>>>> aa9e82687393c774f54c174de9a0f4024543684d
dep_fix("stringr", "1.5.0")
```

This performs a single-package rollback only -- it does not resolve
cascading conflicts the rollback might introduce elsewhere. For full
dependency resolution, use `renv::restore()` or `pak`'s solver.
